---
title: "DS6306 Case Study 1"
author: "Rick Fontenot"
date: "12/20/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

For this case study we assume that our audience is the CEO and CFO of Budweiser (our client) and that they only have had one class in statistics. They have hired us to answer 7 questions and beyond those general questions we will speculate / anticipate what may be of interest to them.

We will start by importing the following data for analysis:

Beers.csv:
Name: Name of the beer.
Beer_ID: Unique identifier of the beer.
ABV: Alcohol by volume of the beer.
IBU: International Bitterness Units of the beer.
Brewery_ID: Brewery id associated with the beer.
Style: Style of the beer.
Ounces: Ounces of beer.

Breweries.csv:
Brew_ID: Unique identifier of the brewery.
Name: Name of the brewery.
City: City where the brewery is located.
State: U.S. State where the brewery is located.

```{r}
library(dplyr)
library(tidyverse)


beers = read.csv("https://raw.githubusercontent.com/rickfontenot/DS6306_Study1/main/Beers.csv", header = TRUE)

breweries = read.csv("https://raw.githubusercontent.com/rickfontenot/DS6306_Study1/main/Breweries.csv", header = TRUE)


head(beers)
head(breweries)

#install.packages("visdat")
library(visdat)
vis_dat(beers, warn_large_data=FALSE)
vis_dat(breweries, warn_large_data=FALSE)

dim(beers) #2410 observations x 7 columns
dim(breweries) #558 observations x 4 columns

sum(is.na(beers$Brewery_id)) #there are no missing values on the merge var
sum(is.na(breweries$Brew_ID)) #there are no missing values on the merge var

beers <- beers %>% rename(Beer_Name = Name) #both dataframes have "Name"
breweries <- breweries %>% rename(Brewery_Name = Name)

beerbrew <- merge(beers,breweries, by.x = "Brewery_id", by.y = "Brew_ID")
head(beerbrew)
 
head(beerbrew,6)
tail(beerbrew,6) # added to complete Question 2

beerbrew["Class"] <- ifelse(str_detect(beerbrew$Style,"IPA"),"IPA", ifelse(str_detect(beerbrew$Style,"Ale"),"Ale", ifelse(str_detect(beerbrew$Style,"Lager"),"Lager", ifelse(str_detect(beerbrew$Style,"Stout"),"Stout", ifelse(str_detect(beerbrew$Style,"Pilsner"),"Pilsner", ifelse(str_detect(beerbrew$Style,"Pilsener"),"Pilsner", ifelse(str_detect(beerbrew$Style,"Porter"),"Porter", ifelse(str_detect(beerbrew$Style,"APA"),"APA", ifelse(str_detect(beerbrew$Style,"Cider"),"Cider", ifelse(str_detect(beerbrew$Style,"Witbier"),"Witbier", ifelse(str_detect(beerbrew$Style,"Kölsch"),"Kölsch", ifelse(str_detect(beerbrew$Style,"Fruit"),"Fruit", ifelse(str_detect(beerbrew$Style,"Hefeweizen"),"Hefeweizen", ifelse(str_detect(beerbrew$Style,"Oktoberfest"),"Oktoberfest", ifelse(str_detect(beerbrew$Style,"Bitter"),"Bitter",
"Other")))))))))))))))

beerbrew %>%ggplot(aes(x = Class)) + geom_bar() +
ggtitle("Distribution of Styles") + coord_flip() + xlab("Style") + ylab("Count")

library(GGally)
beerbrew %>% select(ABV, IBU, Class) %>% ggpairs(aes(color=Class))

beerbrew %>% ggplot(mapping=aes(y=IBU, x=ABV,color=Class)) + geom_point(size=0.5, position="jitter")+ geom_smooth(method="lm",se=FALSE,size=1)+ labs(title="Correlation between ABV and IBU")
```


```{r, Joe's EDA}
#Joe EDA plotting on a map. 
#install.packages('usmap')
library(usmap)
```


```{r, EDA Question 1 How many Brewerues are in Each State?}
#Joe EDA plotting on a map. 
#install.packages('usmap')
library(usmap)

#finding the number of brewers by counting distinct brewery ID for each state
brews.per.st <- beerbrew %>% group_by(State) %>% summarize(num.brews = n_distinct(Brewery_id))

#trimming extra space SHOUT OUT Rick for being a homie and finding this
brews.per.st$State <- trimws(brews.per.st$State)

#needed to make column name lower case to match usmaps library vector of states
brews.per.st <- brews.per.st %>% rename(state = State)

#visualize Number of Breweries Per State
brews.per.st %>% ggplot(aes(x=reorder(state, -num.brews), y=num.brews,color=state)) + 
  geom_bar(stat = "identity", position = "dodge") + theme(plot.title= element_text(hjust = 0.5),axis.text.x=element_text(angle=90,size =   rel(.6),margin = margin(.05, unit = "cm"),vjust =0)) + 
  labs(title = "Number of Breweries by State", y = "Number of Breweries",x= 'State')

#Visualize the Bar Graph on Map of USA
brews.per.st.map <- plot_usmap(data = brews.per.st, values = "num.brews", labels = FALSE, label_color = "white",  color ="black", alpha = 0.8, size = 1) + scale_fill_continuous(low="blue", high = "red",name = "Number of Breweries per State", label = scales::comma) + theme(legend.position = "bottom") +
  ggtitle("Heat map of Breweries per State")

brews.per.st.map
```


```{r, Joe's EDA, Question 3}
#Question 3 Address Missing values in each col

colSums(is.na(beerbrew))

#ABV has 62 missing values
#IBU has 1005 missing values 2410 total 
#no other NAs found
#it is likely that different state differ on their requirement to track IBU and ABV to operate within that state.
```
Impute missing values

```{r}
library(mice)
md.pattern(beerbrew)

library(VIM)
aggr_plot <- aggr(beerbrew, col=c('navyblue','red'), numbers=TRUE, sortVars=TRUE, labels=names(beerbrew), cex.axis=.7, gap=3, ylab=c("Histogram of missing data","Pattern"))

marginplot(beerbrew[c(4,5)])

tempData <- mice(beerbrew,m=5,maxit=50,meth='pmm',seed=500)
xyplot(tempData,ABV ~ IBU,pch=18,cex=1)
densityplot(tempData)

beerbrew_imputed <- complete(tempData,1)
head(beerbrew_imputed)

md.pattern(beerbrew_imputed)

aggr_plot <- aggr(beerbrew_imputed, col=c('navyblue','red'), numbers=TRUE, sortVars=TRUE, labels=names(beerbrew_imputed), cex.axis=.7, gap=3, ylab=c("Histogram of missing data","Pattern"))

#Merge imputed value columns into main data frame
colnames(beerbrew_imputed)[4] = "ABV_imputed" 
colnames(beerbrew_imputed)[5] = "IBU_imputed" 

temp <- beerbrew_imputed %>% select(Beer_ID, ABV_imputed, IBU_imputed)

beerbrew <- left_join(beerbrew, temp,by = "Beer_ID")

beerbrew
```



```{r, Joe's EDA, Compute Median by state ABV and IBU}
#Question 4 compute median median abv and ibu
#i'm making the choice to omit given 1005 or 2410 na values

st.ibu <- aggregate(IBU~State, data =  beerbrew, median, na.action=na.omit)
st.ibu

###visualize the state IBU median
st.ibu %>% ggplot(aes(x=State,y=IBU, fill=IBU)) + scale_fill_distiller(palette='Spectral') + geom_bar(stat = "identity") +
  theme(axis.text.x=element_text(angle=45,size = rel(.8), 
        margin = margin(.05, unit = "cm"),vjust =.5, hjust = 1))


#by date median ABV omitted na values
#Here it is practical to impute the missing values
st.abv <- aggregate(ABV~State, data = beerbrew, median, na.action = na.omit)
st.abv

#visualize the state abv medians
st.abv %>% ggplot(aes(x=State,y=ABV, fill=ABV)) + scale_fill_distiller(palette='Spectral') + geom_bar(stat = "identity") +
  theme(axis.text.x=element_text(angle=45,size = rel(.8), 
        margin = margin(.05, unit = "cm"),vjust =.5, hjust = 1))

#x is a table filtered for the missing abv values.
x <- beerbrew[!complete.cases(beerbrew$ABV),]

```
```{r, Question 5, which state has the highest ABV / IBU}

max.abv <- beerbrew %>% slice_max(ABV) %>% summarise(State, Beer_Name, ABV)
#Answer Colorado, Lee Hill Series Vol.5, ABV = 0.128
max.abv

max.ibu <- beerbrew %>% slice_max(IBU) %>% summarise(State, Beer_Name, IBU)
#Answer Oregon, Bitter Bitch Imperial, IBU = 138
max.ibu
 

```
Re-do State medians with imputed values for comparison

```{r}
state.medians <- beerbrew %>% group_by(State) %>% summarise(ABV.median = median(ABV,na.rm = TRUE),IBU.median = median(IBU,na.rm = TRUE), ABV_imputed.median = median(ABV_imputed),IBU_imputed.median = median(IBU_imputed))

state.medians

p_ABV = ggplot(state.medians, aes(label=State)) + 
  geom_point(aes(ABV.median,ABV_imputed.median))

p_IBU = ggplot(state.medians, aes(label=State)) + 
  geom_point(aes(IBU.median,IBU_imputed.median))

library(plotly)
ggplotly(p_ABV) #DE shows big change in median with imputed, why?
ggplotly(p_IBU) #ME shows big change in median with imputed, why?
```
```{r, Question 6, Comment on Summary statistics distribution of ABV variable}


summary(beerbrew$ABV)

#Distribution of ABV Variable 
beerbrew %>% select(ABV) %>% ggplot(mapping=aes(x= ABV)) +
  geom_histogram(bins = 50) +  labs(title = "Distribution of Alcohol By Volume", y = "Frequency",x= 'ABV')

# Compare by State 
beerbrew %>% select(ABV, State) %>% ggplot(mapping=aes(x=ABV)) +
  geom_histogram(bins = 50) + facet_wrap(~State) + labs(title = "Distribution of Alcohol By Volume", y = "Frequency",x= 'ABV')
  

```

```{r Jason's EDA, Question 6 Start of Density plot by Beer Class}
p <- ggplot(beerbrew, aes(x=ABV)) + 
  geom_density()
p
```



```{r Joe's EDA, Question 7, Is there a relations between IBU and ABV}

beerbrew %>% ggplot(mapping=aes(y=IBU, x=ABV)) + geom_point()+
  geom_point(position=position_jitter(width=0.01),alpha=0.5)+
  geom_smooth(method="lm",se=FALSE,size=2)+
  labs(title="Correlation between ABV and IBU")


# correlation test

IbuAbv.cor <- cor.test(beerbrew$IBU,beerbrew$ABV, method= 'pearson')
IbuAbv.cor # cor test result 0.6706 positive linear correlation

```


```{r Jason's EDA, Question 7, Is there a relations between IBU and ABV}
library(ggpubr)
#7 Scatter Plot
#Feel free to remove

r_test = cor.test(beerbrew$IBU, beerbrew$ABV)
r_test

beerbrew.mod <- lm(IBU ~ ABV, data = beerbrew) 
summary(beerbrew.mod)

ggscatter(beerbrew, x = "IBU", y = "ABV", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          xlab = "IBU", ylab = "ABV")
```

```{r Jason's EDA, Question 8, KNN}
library(caret)
library(e1071)
library(class)
#Start of KNN
#Work in progress

# Filter out any beer style that is not an IPA or an Ale.
beers_IPAALE <- filter(beerbrew,grepl('IPA|Ale',Style))
beers_IPAALE <- beers_IPAALE[,-8]
# Create a new data frame that only holds the ABV,IBU, and beer style.
beers_IPAALE_sub <- select(beers_IPAALE,ABV,IBU,Style)
beers_IPAALE_sub$Style <- as.character(beers_IPAALE_sub$Style)

# Normalize the Beer styles to IPA or ALE
# This loop is used to iterate through each row and then to normalize the styles to simply say IPA or Ale.

for (i in 1:1534) {
  if (is.na(str_match(beers_IPAALE_sub[i,3],".Ale"))) {
    beers_IPAALE_sub[i,3] <- "IPA"
  } else {
    beers_IPAALE_sub[i,3] <- "ALE" 
    
  }
}


set.seed(877)
splitpale = .70
beerbrewsample <- sample(1:dim(beers_IPAALE_sub)[1],round(splitpale * dim(beers_IPAALE_sub)[1]))


# Test and Training with 70/30 split
knn_train <- beers_IPAALE_sub[beerbrewsample,]
knn_test <- beers_IPAALE_sub[-beerbrewsample,]

knn_train = na.omit(knn_train)
knn_test = na.omit(knn_test)

#md.pattern(knn_train)



# Using only the IBU and ABV values from both the training and test sets
# I use the beer style as the class against which the knn will search.
knnbeer <- knn(knn_train[,1:2],knn_test[,1:2],cl=knn_train$Style,k=24,prob = TRUE)
# The confusion matrix is used for calibrating the output of a model and examining all possible outcomes of the predictions
ipamatrix <- confusionMatrix(table(knnbeer,knn_test$Style))
ipamatrix

```

```{r}
beerbrew %>%ggplot(aes(x = Brewery_id)) + geom_bar() +
ggtitle("Distribution of Styles") + coord_flip() + xlab("Brewery ID") + ylab("Count")

brewsize <- beerbrew %>% count(Brewery_id)

beerbrewsize <- merge(brewsize,beerbrew, by.x = "Brewery_id", by.y = "Brewery_id")
head(beerbrewsize)

beerbrewsize %>%ggplot(aes(x = n, color=Class)) + geom_bar() +
ggtitle("Distribution of Styles") + xlab("#Beers per brewery") + ylab("Count")

```


US Map exercise


```{r}
library(maps)
install.packages("mapproj")
library(mapproj)

lookup = data.frame(abb = state.abb, State = state.name) #makes a data frame with State name and abbreviation.

BreweryMapData = breweries %>% count(State) #count up the breweries of each state.

colnames(BreweryMapData)[2] = "num_breweries" #change "n" to "num_breweries"
colnames(BreweryMapData)[1] = "abb" #change for merge
BreweryMapData$abb <- trimws(BreweryMapData$abb)
BreweryMapData <- BreweryMapData %>% filter(abb!='DC')

BreweryMapData = merge(BreweryMapData,lookup,"abb")
colnames(BreweryMapData)[3] = "region" #change for merge
BreweryMapData$region <- tolower(BreweryMapData$region)

states <- map_data("state")

map.df <- merge(states,BreweryMapData, by="region", all.x=T)
map.df <- map.df[order(map.df$order),]

ggplot(map.df, aes(x=long,y=lat,group=group))+
  geom_polygon(aes(fill=num_breweries))+
  geom_path()+ 
  scale_fill_gradientn(colours=rev(heat.colors(10)),na.value="grey90")+ggtitle("Number of Breweries by State")+
coord_map()

```

```{r, Question 9, Brewery Strength Index}
#comparing the number of breweries to the per capita alcohol consumption by taking per capita consumption / number of breweries in each state. I'm calling this ration a brewery strength index. It might help reveal which states are under presented in terms of breweries.

per.capita <-  read.csv("https://github.com/rickfontenot/DS6306_Study1/raw/main/NIH_per_capita.csv", header = TRUE)
head(per.capita)

#change to state abbreviations
per.capita$State <- state.abb[match(per.capita$State, state.name)]

per.capita$State[which(is.na(per.capita$State))] <- "DC"

#modify beerbrew IOT summarize number of breweries by state
brews.per.st <- beerbrew %>% group_by(State) %>% summarize(num.brews = n_distinct(Brewery_id))

#trimming extra space SHOUT OUT Rick for being a homey and finding this
brews.per.st$State <- trimws(brews.per.st$State)


#join two data to compare per capita consumption to state breweries
brew.consum <- full_join(brews.per.st, per.capita, by= "State", Copy= TRUE)


#create ratio of gallons per capita to number of breweries

brew.consum <- brew.consum %>% mutate(per.capita.brewery.strength = 
                                        Gallons.of.Ethanol.Per.Capita / num.brews) 

#create rankings based on the ration above
brew.consum <- brew.consum %>%
  mutate(per.capita.rank = dense_rank((desc(per.capita.brewery.strength))))

#ordering states for preparation for plotting
brew.consum <- brew.consum[order(brew.consum$State), ]

#needed to make column name lower case to mat usmaps library vector of states
brew.consum <- brew.consum %>% rename(state = State)

###density plot. May not be informative probably delete commented it out ###
#brew.consum %>% ggplot(aes(x=per.capita.brewery.strength)) +geom_density() 


#makeing a bar chart to help show which states have the most advantageous ratio
####fixed### bar chart sorting by a different method goal is to remove the other coloumn
top10.per.capita <- brew.consum %>% arrange(desc(per.capita.brewery.strength)) %>% slice(1:10)

#bar chart of top 10 most advantagous ratio. Removed Other coloumn
top10.per.capita %>% ggplot(aes(x=reorder(state, -per.capita.brewery.strength), y = per.capita.brewery.strength)) +
  geom_bar(aes(fill= state),stat = "identity") +
  labs(title = "Top 10 States. Number of breweries compared to per capita consumption", 
       x="Top 10 States", y="Per capita consumption / Num of Breweries")



library(usmap)

ratio.map <- plot_usmap(data = brew.consum, values = "per.capita.brewery.strength",
                        labels = FALSE, label_color = "white",  color ="black", alpha = 0.8,
                        size = 1) + 
  scale_fill_continuous(low="blue", high = "red",
                        name = "Breweries to Per capita consumption",
                        label = scales::comma) + theme(legend.position = "bottom") +
  ggtitle("Under Represented States")

ratio.map

###        if we want state lables; set labels = TRUE 
#        and runs this code below for smaller labels. Helps with NE region  ###
#set label font size
#  ratio.map$layers[[2]]$aes_params$size <-2
#  print(ratio.map)
```
```{r, 2019 census data combined in }

#2019 Census data
census2019 <-  read.csv("https://raw.githubusercontent.com/rickfontenot/DS6306_Study1/main/nst-est2019.csv", header = TRUE)

#data preparation removing top rows for regional grand total US pop
census2019 <- census2019[-c(1,2,3), ]

#data prep renaming coloumn title to State
census2019 <- census2019 %>% rename('State' = "ï..table.with.row.headers.in.column.A.and.column.headers.in.rows.3.through.4...leading.dots.indicate.sub.parts.")

#data prep renaming coloumn for 2019 state population to pop.2019
census2019 <- census2019 %>% rename("pop.2019" = "X.11")

#data prep selecting only state and corresponding population columns
census2019 <- subset(census2019, select = c("State", "pop.2019"))

#data prep using regex to remove "." before state names
census2019$State <- str_replace_all(census2019$State, "^\\.", "")

#data prep re ordering rows for join
rownames(census2019) <- 1:nrow(census2019)

# data prep more rows needed to be removed for the join
census2019 <- census2019[-c(1:5, 57:63), ]
head(census2019)
#data prep re ordering rows for the join
rownames(census2019) <- 1:nrow(census2019)

#data prep function for removing commas in numbers and setting values to numeric
replaceCommas<-function(x){
  x<-as.numeric(gsub("\\,", "", x))
}

#data prep implementing comma replace function
census2019$pop.2019 <- replaceCommas(census2019$pop.2019)

#data prep matching full state name to state abbrievations
census2019$State <- state.abb[match(census2019$State, state.name)]

#data prep DC isn't part of the vector manually entering it here by sorting by NA was only NA value
census2019$State[which(is.na(census2019$State))] <- "DC"

#data prep renaming column name to match to lower case "s"
census2019 <- census2019 %>% rename("state" = "State")

#data prep merge
brew.consum <- full_join(brew.consum, census2019, by = 'state')

#applying brewery strength index to total population
brew.consum <- brew.consum %>% mutate(pop.adjusted.brewery.strength = 
                         pop.2019 * per.capita.brewery.strength) 


#visualize top 10 population adjusted brewery stength
#slicing the top 10 states with the highest pop adjusted brewery strength index
top10 <- brew.consum %>% arrange(desc(pop.adjusted.brewery.strength)) %>% slice(1:10)

#map heat map of US for brewery strength index applied to total state pop
pop.adjusted <- plot_usmap(data = brew.consum, values = "pop.adjusted.brewery.strength",
           labels = FALSE, label_color = "white",  color ="black", alpha = 0.8,
           size = 1) + 
  scale_fill_continuous(low="blue", high = "red",
  name = "Population adjusted Brewery Strength",
  label = scales::comma) + theme(legend.position = "bottom") +
  ggtitle("Undere Represented States Adjusted for pop size")

pop.adjusted

top10 %>% ggplot(aes(x=reorder(state, -pop.adjusted.brewery.strength), y = pop.adjusted.brewery.strength)) +
  geom_bar(aes(fill= state),stat = "identity") + labs(title = "Population Adjusted Brewery Strength", x="Top 10 States", y = "State Brewery Strength * Total Pop" )

#These states represent what might be the under developed markets in terms of craft breweries. 

```